---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# composits

<!-- badges: start -->
<!-- badges: end -->

The goal of *composits* is to find outliers in compositional, multivariate and univariate time series. It is an ensemble method that uses the packages *forecast*, *tsoutliers*, *anomalize* and *otsad*. 

## Installation

<!-- You can install the released version of composits from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("composits") -->
<!-- ``` -->
As this is a private repo you need to generate a token (PAT) to install this using *install_github*. This is explained at http://devtools.r-lib.org/reference/install_github.
And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sevvandi/composits", auth_token ='yourtoken')
```

Or you can do the following on your RStudio Terminal after you pull the repo. 

```{r installing, eval=FALSE}
cd ..
R CMD build composits
R CMD INSTALL composits_0.0.0.9000.tar.gz
```

## Example

```{r example}
library(composits)
set.seed(100)
n <- 600
x <- sample(1:100, n, replace=TRUE)
x[25] <- 200
x[320] <- 300
x2 <- sample(1:100, n, replace=TRUE)
x3 <- sample(1:100, n, replace=TRUE)
X <- cbind.data.frame(x, x2, x3)
x4 <- sample(1:100, n, replace=TRUE)
X <- cbind.data.frame(x, x2, x3, x4)
out <- mv_tsout_ens(X)
out$all
out$outliers
```

